CODECHECK: An open-science initiative to facilitate sharing of computer programs and results presented in scientific publications.

### 2020-12-04 codecheck-paper.tex
add todo
try to fix namerefs
add "main contribution" statement

\documentclass[12pt]{article}
\usepackage{f1000_styles}
\usepackage[numbers]{natbib}
\usepackage{hyperref,url,setspace}

\usepackage{booktabs}
\usepackage{doi} % turn DOIs into links

@@

Given that there are no clear standards yet for sharing data, what hope
might there be for sharing computer programs? Our anectodal experiences
on this matter suggest that often a variety of reasons are given for why
their code cannot be shared, e.g.~``there is no documentation / I do not
want to maintain it / I do not want to give away my code to my
competitors''. Our view, rooted in open science princples, is that
wherever possible sharing code is good for the community, as outlined
ten years ago \cite{Barnes2010-iv}. Having the code freely available, and
suitably archived, provides a valuable resource for others to learn
from, even when it doesn't run or if there is no documentation. However,
with a little effort, we believe that if someone independent can run/execute
the code, this is worth documenting as early as possible. With this in
mind, we have developed a set of PRINCIPLES and an example workflow that
provides what he hope is a pragmatic way of checking that code works. We
call this system CODECHECK as outlined next.

@@

The time for a CODECHECK greatly depends on the intentions and the ascribed
importance. The earlier a CODECHECK happens in the process, the more the
results can play a role in the most important decision in scholarly 
communication: is a paper published, sent back for revisions, or rejected.
The later in a review process the check happens, the less challenging is it
to allow a bidirectional communication between author and codechecker, e.g.,
because the author might already be notified of the acceptance and may more
freely share materials online.
Even earlier checks, i.e., a CODECHECK of a preprint, may even help to 
improve the workflow itself even before a publisher is involved. The 
possibility for codechecking papers could be part of a preprint server's
policy, or simply be initiated by interested readers and community members.

>>>>

\documentclass[12pt]{article}
\usepackage{f1000_styles}
\usepackage[numbers]{natbib}
\usepackage{hyperref}
\usepackage{url}
\usepackage{setspace}

\usepackage{booktabs}
\usepackage{doi} % turn DOIs into links
\usepackage{nameref}

@@

Given that there are no clear standards yet for sharing data, what hope
might there be for sharing computer programs? While both data and software
together are required to validate outcomes of a computational analysis,
the do differ in the respect that the former can be seen as static/inert,
while the latter requires an environment and interaction. This makes 
software harder to share.
Our anectodal experiences
on this matter suggest that often a variety of reasons are given for why
code cannot be shared, e.g.~``there is no documentation / I do not
want to maintain it / I do not want to give away my code to my
competitors''. Our view, rooted in Open Science princples, is that
wherever possible sharing code is good for the community, as outlined
ten years ago \cite{Barnes2010-iv}. Having the code freely available, and
suitably archived, provides a valuable resource for others to learn
from, even when it doesn't run or if there is no documentation. However,
with a little effort, we believe that if someone independent can run/execute
the code, this is worth documenting as early as possible. With this in
mind, we have developed a set of PRINCIPLES and an example workflow that
provides what he hope is a pragmatic way of checking that code works.

The main contribution of this work is a thorough description of a process
and its variations to integrate a much needed evaluation of the 
computational reproducibility \cite{barba_terminologies_2018} into peer
review, and a demonstration of its feasibility by means of over 20
reproductions across different scientific disciplines.
We call this system CODECHECK as outlined next.

@@

The time for a CODECHECK greatly depends on the intentions and the ascribed
importance. The earlier a CODECHECK happens in the process, the more the
results can play a role in the most important decision in scholarly 
communication: is a paper published, sent back for revisions, or rejected.
The later in a review process the check happens, the less challenging is it
to allow a bidirectional communication between author and codechecker, e.g.,
because the author might already be notified of the acceptance and may more
freely share materials online.
Even earlier checks, i.e., a CODECHECK of a preprint, may even help to 
improve the workflow itself even before a publisher is involved. The 
possibility for codechecking papers could be part of a preprint server's
policy, or simply be initiated by interested readers and community members.
% TODO add a sentence about the possible effect of just asking for
% code/data early on in the submission process, see https://twitter.com/tsuyomiyakawa/status/1230672758163402752?s=09)
